Eagle: Making multiple-locus association mapping on a genome-wide scale routine

### 2018-07-26 paper.tex
27/7 refinement v1

We were interested in answering the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations.  To 
answer this question, we calculated, empirically, 
the statistical power and false discovery rates of Eagle and the other methods across the six scenarios
For each replicate, we knew which SNPs  had been used to create the quantitative trait data. These SNPs are in true association with 
the trait. By knowing these SNPs, we were able to assess the validity of a  method's findings.

The power (false discovery rate) of a method, under a particular scenario,  was calculated by taking the median of its power (false discovery rate) 
values across the 100 replicates of the scenario. To calculate the power of a method from its analysis of a replicate, we divided the number of findings for that method that were true by the number of SNP in true association with the trait. To calculate the false discovery rate of a method from its analysis of a replicate, 
we instead divided by the number of  findings for that method. 
A finding was counted as true if the SNP was located within 40 kilobase pairs of a SNP in true 
association with the trait.

The power and false discovery rate of Eagle and the other multiple-locus methods across the scenarios 150 x 5K, 350 x 500K, 1500 x 50K, and 2000 x 500K are shown in Figure \ref{figpowermultiple}.  We restricted our attention to these scenarios for the multiple-locus methods because for scenario 4000 x 1.5M, 
the data could only be analysed by Eagle, MLMM, and r2VIM. For scenario 10000 x 1.5M, the data could only be analysed by Eagle. 
The power and false discovery rate of Eagle and the two single-locus methods, GEMMA and FaST-LMM,  are shown in
Supplementary Figure \ref{supfigpowersingle}.  Each plot contains 
single points and power curves. The single points are the power and false discovery rates for Eagle and MLMM.
These two methods treat association mapping as a model selection problem. Their are no significance thresholds to be set. 
The power curves are for those methods that treat association mapping as a parameter estimation problem. Here, the 
significance of the findings are assessed against a significance threshold. The power curves in the plot show how power changes with 
the false discovery rate as the significance threshold  is adjusted.

>>>>

We answer the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations. We do this by estimating the 
power and false discovery rates of Eagle and the other methods for the six scenarios.  Since, for a replicate, we knew which SNP were assigned additive effects in 
creating its trait data, we knew the SNP that were in true association with the trait. This allowed us to assess if a method's findings were true or 
false. A  finding was counted as true if it was positioned within 40 kilobase pairs of the location of a SNP in true association. 
When a replicate was analysed, we obtained an estimate of power, the number of findings that were true for the method divided by the 
number of SNPs in true association. We also obtained an estimate of the false discovery rate, the number of findings that were false for the method 
divided by the number of (true and false) findings for the method.  Both these estimates varied with replicate. 
The power (false discovery rate) of a method, for a scenario, was found by taking the median of the power (false discovery rate) estimates that were calculated from the analyses of 
the 100 replicates.

The power and false discovery rates of Eagle and the other multiple-locus methods across the scenarios 150 x 5K, 350 x 500K, 1500 x 50K, and 2000 x 500K are shown in Figure \ref{figpowermultiple}.  We restricted our attention to these scenarios for the multiple-locus methods because for scenario 4000 x 1.5M, 
the data could only be analysed by Eagle, MLMM, and r2VIM. For scenario 10000 x 1.5M, the data could only be analysed by Eagle. 
The power and false discovery rate of Eagle and the two single-locus methods, GEMMA and FaST-LMM,  are shown in
Supplementary Figure \ref{supfigpowersingle}.  Each plot contains 
single points and power curves. The single points are the power and false discovery rates for Eagle and MLMM.
These two methods treat association mapping as a model selection problem. Their are no significance thresholds to be set. 
The power curves are for those methods that treat association mapping as a variable selection problem. Here, the 
significance of the findings are assessed against a significance threshold. The power curves in the plot show how power changes with 
the false discovery rate as the significance threshold  is adjusted.
